The seasonal cycle of upper-ocean mixing at 8°N in the Bay of Bengal

### 2019-09-25 paper1.tex
Updates from Overleaf

\abstract{
We describe the seasonal cycle of mixing in the top \SIrange{30}{100}{m} of the Bay of Bengal as observed by moored mixing meters (χpods) deployed along 8°N between 85.5°E and 88.5°E in 2014 and 2015.
All χpod observations were combined to form seasonal-mean vertical profiles of turbulence diffusivity $K_T$ in the top 100m.
The strongest turbulece is observed between July and November during the southwest (May -- September) and post-monsoon seasons (October -- November).
The northeast (December -- February) monsoon is a period of similarly high mean diffusivities but an order of magnitude lower median $K_T$, a sign of energetic episodic mixing events forced by near-inertial shear events.
The months of March and April, a period of weak wind forcing and low near-inertial shear amplitude, are characterized by near-laminar flow and near-molecular values of $K_T$ in the thermocline for weeks at a time.
High mixing events coincide with the passage of surface-forced downward-propagating near-inertial waves and with the presence of enhanced low-frequency shear associated with the Summer Monsoon Current between July and October.
We find that monthly-averaged turbulent transport of salt out of the salty Arabian Sea water between August and January is significant relative to local $E-P$.
The magnitude of this salt flux is approximately that required to close model-based salt budgets for the upper Bay of Bengal.
}
\maketitle

@@

The Bay of Bengal (the Bay) is the eastern semi-enclosed basin of the north Indian Ocean and is known to influence large scale atmospheric circulation throughout the tropics (\textbf{CITATION NEEDED}).
The Bay's physical oceanography is characterized by two major features.
First, its circulation reverses seasonally under the influence of the Indian Ocean monsoon --- the seasonal reversal of winds north of approximately 10°S in the Indian Ocean basin; and
Second, it receives an immense amount of freshwater --- more than 50\% of the freshwater runoff into the entire tropical Indian ocean \citep{Sengupta2006,Gordon2016}.
Much of central India's annual rainfall results from convective systems that originate in the Bay between June and September, and then propagate northwestward over the Indian subcontinent \citep{Gadgil2003a}.
Since interannual variations in mean rainfall are strongly correlated with fluctuations in India's agricultural output \citep{Gadgil2006}, understanding air-sea interaction in the Bay is of prime social relevance.
Building such understanding however is complicated by the immense amount of fresh water that enters the Bay every year.
This fresh water is advected and stirred around by Ekman currents as well as the Bay's mesoscale eddy field, forming a thin \(\mO(\SIrange{5}{10}{m})\) cap over the Bay's subsurface waters that heavily influences air-sea interaction in the region \citep{Sengupta2016,SreeLekha2018,Rao2011} (\textbf{NEED BETTER CITATION FOR AIR-SEA INTERACTION}).
Much effort has been directed recently toward understanding the fate and consequences of this fresh layer in the Bay (CITATIONS).

The monsoon and its associated precipitation is visualized using seasonal mean wind stress from the Tropflux estimate \citep{Kumar2012} and precipitation from the TRMM Multi-satellite Precipitation Analysis dataset \citep{trmm} in Figure \ref{fig:ppt}.
Between May and September (southwest or SW monsoon), strong southwesterly winds bring substantial precipitation to the Indian subcontinent (Figure \ref{fig:ppt}c).
Over the ocean, the precipitation maximum is over the northeastern Bay (Figure \ref{fig:ppt}c).
The months of October and November (post-monsoon period, SWNE) are characterized by episodic strong basin-scale cyclones that interrupt periods of weak winds --- the basin-wide mean wind stress is small (Figure \ref{fig:spatial}f--j).
Sri Lanka in particular receives substantial precipitation during this season.
Between December and February (northeast or NE monsoon), the winds, now northeasterly, strengthen but are weaker than during the SW monsoon.
Both India and Sri Lanka receive relatively little rainfail during this period with the precipitation maximum located in the Bay.
The months of March and April are a period of weak winds and almost no precipitation north of 4°N (northeast-southwest transition, NESW).

The monsoonal cycle of wind forcing described above imprints seasonality on the Bay's circulation (our Figure \ref{fig:spatial}a--e, \citealp{Schott2002,Shankar2002}).
The East India Coastal Current (EICC) spins up at the Bay's western boundary during both monsoons, flowing northward between May and October and then southward between December and April.
The EICC is readily visible in seasonally-averaged estimates of near-surface ocean velocity from the OSCAR product --- a diagnostic estimate at five-day frequency that ignores local acceleration and nonlinearities but accounts for geostrophic, thermal wind and Ekman currents (vectors in Figure \ref{fig:spatial}a--e; \citealp{Bonjean2002}).
The EICC exists as a discontinuous flow with many recirculation loops \citep{Durand2009} and appears as a region of high mesoscale kinetic energy in maps of geostrophic eddy kinetic energy \(\text{EKE} = 0.5[u_g^2 + v_g^2]\) (Figure \ref{fig:spatial}a--e, color).
Here \((u_g, v_g)\) are precomputed geostrophic velocities provided by the Copernicus Marine Environment Monitoring Service (CMEMS).
They are estimated from delayed-time sea surface height estimates as measured by multiple satellite altimeters.
Apart from the EICC, a strong mesoscale cyclone develops to the east of Sri Lanka during the SW monsoon, termed the Sri Lanka Dome.
In the southwestern Bay, the Summer Monsoon Current (SMC) threads a pathway between the Sri Lanka Dome to the north and an anticyclone to the south, transporting warm salty water in to the Bay.
It is visible as an eastward jet along 8°N between 85°E and 92°E in Figure \ref{fig:spatial}d.
Peak velocity in the SMC can exceed \SI{1.5}{\meter\per\second} and northward transport has been estimated to be in the range \SIrange{10}{27}{Sv} \citep{Wijesekera2016,Vinayachandran1999,Webber2018}.
Meanders in the Summer Monsoon Current forced by Rossby waves that trace their origin back to equatorial wind fluctuations make the southwestern and south-central Bay a site of strong mesoscale activity during the southwest monsoon (our Figure \ref{fig:spatial}a--e, \citealp{Vinayachandran1998,Chen2018}).

@@

The western semi-enclosed basin of the Indian Ocean, the Arabian Sea, is generally considered the source of the required salty water \citep[for example, ][]{Jensen2001}.
Recently \cite{Sanchez-Franks2019} used a multi-year model simulation to argue that the ultimate source of the salty water is the western equatorial Indian Ocean.
Regardless of ultimate source, both observations and models agree that the Summer Monsoon Current (SMC) is the dominant pathway for salty water entering the Bay \citep{Jensen2001,Vinayachandran2013,Webber2018}\footnote{Recent observations and model simulations describe a second pathway — a persistent subsurface inflow of salty water during the NE monsoon that exists as a superposition of frequent salty intrusion events that average out to a region of broad northward flow of high salinity water west of 85°E \citep{Wijesekera2015,Jensen2016}.}.
The salty signature of the SMC is visible in maps of the depth of the 34.75 isohaline surface which shallows by \SI{25}{m} or so in the southwestern Bay during the summer monsoon (Figure \ref{fig:spatial}k--o; \citealp{Murty1992a,Vinayachandran2013}).
The shallow depth of the \(S = \SI{35}{psu}\) isohaline in the southwestern and south-central Bay relative to the northern Bay led \cite{Vinayachandran2013} to hypothesize that the southern Bay is a site of enhanced mixing and upward salt fluxes that may be an important contributor to the salt budget of the Bay.
In agreement with this hypothesis, model studies have implicated vertical mixing as the primary mechanism for diluting the immense amount of fresh water the Bay receives during the southwest monsoon \citep{Akhil2014,Benshila2014,Wilson2016a}.

@@

Figure \ref{fig:nrl} presents the seasonal cycle of winds, turbulence, shear and stratification at mooring NRL5 using daily averaged quantities.
We choose to highlight mooring NRL5  for two reasons.
First, it experiences the least blowdown (\SIrange{10}{20}{\meter}, Figure \ref{fig:nrl}f) and is least contaminated by the associated space-time aliasing.
Second, the turbulence quantities in Figure \ref{fig:nrl} are inferred from measurements recorded by the deep χpod at 105m.
This instrument is the deepest deployed in the Bay to date, and recorded the longest period of
near-molecular diffusivity observed during the transition months of March and April.
The filtered shear components shown in Figure \ref{fig:nrl}d are obtained by first subsampling the filtered depth-time fields along the χpods trajectory and then normalizing by 30-day lowpass filtered \(N²\).
Time series recorded at the other moorings are presented in the Supplementary Material.

@@

We now synthesize Figures \ref{fig:vert}, \ref{fig:nrl} and \ref{fig:shears} to describe the seasonal cycle of shear and turbulence.

@@

Our most dramatic observation is that the χpod at 104-m recorded near-laminar flow in the thermocline with near-molecular values of \(K_T\) during the \emph{entire} month of April.
Similar periods of weak to no mixing are visible at the other χpods but for shorter periods of time (\textbf{Do SOMETHING HERE}).
This is a period of weak winds, weak currents, high net surface heat flux and low near-inertial energy flux (Figures \ref{fig:nrl} and \ref{fig:spatial}).
Similar periods of low to negligible mixing are present at other χpods --- \emph{median}  \(K_T ≤ \SI{1e-6}{\m\squared\per\s} \approxeq 5κ_T\) in most thermocline density bins (deeper distributions in Figure \ref{fig:vert}).
Consistent with the observations of low mixing, the transition months of March and April are a period of weak shear in the thermocline (Figure \ref{fig:shears}). \textbf{COMMENT ON N²}
Weak pulses of near-inertial shear are seen in Figures \ref{fig:shears} and \ref{fig:nrl}e; again this is consistent with weak wind forcing at the surface (Figure \ref{fig:spatial}k--o).

@@

Elevated mixing at NRL5 occasionally lines up with short periods of elevated low frequency shear, indicating presence of the SMC, between May and October (Figure \ref{fig:nrl}e)
The SMC is visible in \(S\low\) at all three moorings during this season though for differing lengths of time (Figure \ref{fig:shears}).
Both seasonal mean surface velocities from the OSCAR product and mooring ADCP data show the SMC to be prominent especially at NRL3 and NRL4, the two westernmost moorings along 8°N (also see Figures \ref{fig:spatial}a--e, \ref{fig:hov} and \citealp{Wijesekera2016a}).
This inference is consistent with the ADCP measurements (Figure \ref{fig:shears})

@@

There is a strong seasonal cycle in thermocline mixing (Figure \ref{fig:vert}) that appears to be linked to a seasonal cycle in thermocline shear (Figure \ref{fig:shears}).
The seasonal cycle in shear results from
(a) the seasonal presence of the Summer Monsoon Current which greatly increases low-frequency shear \(S\low\) between July and October, and
(b) episodic energetic downward propagating near-inertial waves observed outside March and April.
At times, \(S\low\) is of comparable magnitude to near-inertial shear \(S\niwp\) (Figure \ref{fig:shears}).
The seasonal cycle in low-frequency shear is expected from the well established seasonal spinup and spindown of the SMC \citep{Schott2001a}.
A seasonal cycle in near-inertial shear is perhaps expected from the seasonal cycle of winds but our \emph{in-situ} record is not long enough to properly characterize the magnitude of the seasonal cycle in near-inertial energy.
Next we examine the consequences of turbulent mixing in the thermocline.

@@

Is observed seasonally-enhanced mixing in the south-central Bay's thermocline between May and November
important for the Bay's salt budget?
The climatological depth of the \(S=\SI{34.75}{psu}\) surface at 8°N estimated using the Argo mapped climatology shallows by \SI{20}{m} or so between May and November relative to other months (Figures \ref{fig:spatial}k--o and \ref{fig:map}f--i).
The climatological depth of this isohaline is visualized relative to the seasonal variation of thermocline diffusivity by the thick orange horizontal line in Figure \ref{fig:vert}.
Mean diffusivity at this isohaline is approximately \(\SI{1e-4}{\m\squared\per\s}\) during the SW monsoon and the post-monsoon period (SWNE) i.e. between May and November.
Diffusivity is an order of magnitude lower during the NE monsoon and near-molecular during the NESW transition.
Since seasonally averaged surface velocities show the mean path of the SMC to be along the mooring line at 8°N (NRL3, NRL4, and NRL5; Figure \ref{fig:spatial}a--e), we now attempt to quantify turbulent salt flux along 8°N in the south-central Bay using our admittedly sparse dataset.

@@

Due to mooring blowdown we are likely underestimating the true magnitude of turbulent salt flux.
For example, all χpods at 8°N are forced down approximately \SI{50}{m} or so by the Summer Monsoon Current in July during which time they record little turbulent salt flux (Figure \ref{fig:8njs}a).
Inspection of the velocity fields shows that the χpods dive beneath the region of greatest shear in the water column and are likely missing the regions of greatest mixing during this period (Supplementary Information, Figures SXXXXXX \textbf{TODO}).
Given these uncertainties, we do not consider Figure \ref{fig:8njs}c a good estimate of the amplitude of the seasonal cycle of turbulent heat flux but instead interpret it as evidence that climatologically important turbulent fluxes occur in the south-central Bay at least between August and January.

@@

Year-long observations of turbulence from a moored mixing meter χpod at three locations along 8°N revealed a seasonal cycle in upper-ocean turbulence in the Bay of Bengal (Figures \ref{fig:map}, \ref{fig:vert} and Table \ref{tab:pods}).
The seasonal cycle of thermocline turbulence is influenced by the winds both through downward propagating near-inertial waves and by the strongly sheared Summer Monsoon Current (Figures \ref{fig:nrl}, \ref{fig:shears} and \ref{fig:nrl5-niw}).
Multiple χpods record extended periods of weak mixing (1--10 \(κ_T\)) between 50m and 100m depth during the months of March and April --- a period of weak winds, weak currents, low near-inertial energy input, weak near-inertial shear and weak low-frequency shear (Figures \ref{fig:spatial}, \ref{fig:shears} and \ref{fig:nrl}; Tables B1,B2).
Despite the extended periods of low mixing, it has been hypothesized that turbulence in the Bay is necessary to close both heat and salt budgets \citep{Shenoi2002,Vinayachandran2013,Wilson2016a}.
Our observations suggest that turbulent salt fluxes of the right magnitude are indeed occurring in the south-central Bay (Section \ref{sec:results}\ref{sec:js}).

@@

Our observations of enhanced thermocline mixing coincide with bursts of near-inertial shear outside the SW monsoon time period.
Characterizing the seasonal cycle of the near-inertial shear field and understanding the mechanisms that drive that seasonal cycle at depth is thus of prime importance.
These mechanisms include the interaction of near-inertial energy with lower-frequency mesoscale features \citep{Johnston2016} as well as the sensitivity of downward radiation to multi-layer stratification \citep{Lucas2016}.
Another puzzle is the extended period of low mixing during March and April — is the internal wave field much weaker than that expected from the Garrett-Munk spectrum and if so, why?
The Bay offers intriguing opportunities for studying the ocean's internal wave field and its links to turbulence.

>>>>

\abstract{
We describe the seasonal cycle of mixing in the top \SIrange{30}{100}{m} of the Bay of Bengal as observed by moored mixing meters (χpods) deployed along 8°N between 85.5°E and 88.5°E in 2014 and 2015.
All χpod observations were combined to form seasonal-mean vertical profiles of turbulence diffusivity $K_T$ in the top 100m.
The strongest turbulence is observed between July and November during the southwest (May -- September) and post-monsoon seasons (October -- November).
The northeast (December -- February) monsoon is a period of similarly high mean diffusivities but an order of magnitude lower median $K_T$, a sign of energetic episodic mixing events forced by near-inertial shear events.
The months of March and April, a period of weak wind forcing and low near-inertial shear amplitude, are characterized by near-laminar flow and near-molecular values of $K_T$ in the thermocline for weeks at a time.
High mixing events coincide with the passage of surface-forced downward-propagating near-inertial waves and with the presence of enhanced low-frequency shear associated with the Summer Monsoon Current between July and October.
We find that monthly-averaged turbulent transport of salt out of the salty Arabian Sea water between August and January is significant relative to local $E-P$.
The magnitude of this salt flux is approximately that required to close model-based salt budgets for the upper Bay of Bengal.
}
\maketitle

@@

The Bay of Bengal (the Bay) is the eastern semi-enclosed basin of the north Indian Ocean.
It is known to influence large scale atmospheric circulation throughout the tropics (\textbf{CITATION NEEDED}) and has a particularly strong influence on precipitation over the Indian subcontinent.
Much of central India's annual rainfall results from convective systems that originate in the Bay between June and September, and then propagate northwestward over the Indian subcontinent \citep{Gadgil2003a}.
Interannual variations in mean rainfall are strongly correlated with fluctuations in India's agricultural output \citep{Gadgil2006}, lending significant social relevance to the problem of understanding air-sea interaction and near-surface ocean dynamics in the Bay.

%Building such understanding however is complicated by the immense amount of fresh water that enters the Bay every year.
%This fresh water is advected and stirred around by Ekman currents as well as the Bay's mesoscale eddy field, forming a thin \mO(\SIrange{5}{10}{m}) cap over the Bay's subsurface waters that heavily influences air-sea interaction \citep{Sengupta2016,SreeLekha2018,Rao2011}.
%This freshwater cap isolates subsurface waters from surface forcing (for example, \citealp{Lucas2016}). (\textbf{NEED BETTER CITATION FOR AIR-SEA INTERACTION}).
% Much effort has been directed recently toward understanding the fate and consequences of this fresh layer in the Bay (CITATIONS).

The Bay's physical oceanography is characterized by two major features.
First, the Bay's circulation reverses seasonally under the influence of the Indian Ocean monsoon --- the seasonal reversal of winds north of approximately 10°S in the Indian Ocean basin.
Second, the Bay receives an immense amount of freshwater --- more than 50\% of the freshwater runoff into the entire tropical Indian ocean \citep{Sengupta2006,Gordon2016}.
The Indian Ocean monsoon and its associated precipitation is visualized using seasonal mean wind stress from the Tropflux estimate \citep{Kumar2012} and precipitation from the Tropical Rainfall Measuring Mission (TRMM) Multi-satellite Precipitation Analysis dataset \citep{trmm} in Figure \ref{fig:ppt}.
Between May and September (southwest or SW monsoon), strong southwesterly winds bring substantial precipitation to the Indian subcontinent (Figure \ref{fig:ppt}c).
Over the ocean, the precipitation maximum is over the northeastern Bay (Figure \ref{fig:ppt}c).
October and November (post-monsoon period, SWNE) are characterized by episodic strong basin-scale cyclones that interrupt periods of weak winds --- the basin-wide mean wind stress is small (Figure \ref{fig:spatial}f--j).
Sri Lanka in particular receives substantial precipitation during this season.
Between December and February (northeast or NE monsoon), the winds strengthen and switch to being northeasterly, but are weaker than that during the SW monsoon.
Both India and Sri Lanka receive relatively little rainfail during this period with the precipitation maximum located over the Bay.
The months of March and April are a period of weak winds and almost no precipitation north of 4°N (northeast-southwest transition, NESW).

The monsoonal cycle of wind forcing described above imprints seasonality on the Bay's circulation (our Figure \ref{fig:spatial}a--e, \citealp{Schott2002,Shankar2002}).
The East India Coastal Current (EICC) spins up at the Bay's western boundary during both monsoons, flowing northward between May and October and then southward between December and April.
The EICC is readily visible in seasonally-averaged estimates of near-surface ocean velocity from the Ocean Surface Current Analysis Real-time (OSCAR) product.
OSCAR is a diagnostic estimate at five-day frequency that ignores local acceleration and nonlinearities but accounts for geostrophic, thermal wind and Ekman currents (vectors in Figure \ref{fig:spatial}a--e; \citealp{Bonjean2002}).
The EICC exists as a discontinuous flow with many recirculation loops \citep{Durand2009} and appears as a region of high mesoscale kinetic energy in maps of geostrophic eddy kinetic energy \(\text{EKE} = 0.5[u_g^2 + v_g^2]\) (Figure \ref{fig:spatial}a--e, color).
Here \((u_g, v_g)\) are geostrophic velocity anomalies computed from delayed-time sea surface height estimates as measured by multiple satellite altimeters by the Copernicus Marine Environment Monitoring Service (CMEMS)\footnote{DT all-sat-merged Global Ocean Gridded SSALTO/DUACS Sea Surface Height L4 product and derived variables (dataset-duacs-rep-global-merged-allsat-phy-l4-v3).}.
Apart from the EICC, a strong mesoscale cyclone develops to the east of Sri Lanka during the SW monsoon, termed the Sri Lanka Dome.
In the southwestern Bay, the Summer Monsoon Current (SMC) threads a pathway between the Sri Lanka Dome to the north and an anticyclone to the south, transporting warm salty water in to the Bay.
It is visible as an eastward jet along 8°N between 85°E and 92°E in Figure \ref{fig:spatial}d.
Peak velocity in the SMC can exceed \SI{1.5}{\meter\per\second} and northward transport has been estimated to be in the range \SIrange{10}{27}{Sv} \citep{Wijesekera2016,Vinayachandran1999,Webber2018}.
Meanders in the Summer Monsoon Current make the southwestern and south-central Bay a site of energetic mesoscale variability during the SW monsoon (our Figure \ref{fig:spatial}a--e, \citealp{Vinayachandran1998,Chen2018}).

@@

The western half of the north Indian Ocean, the Arabian Sea, is generally considered the source of the required salty water \citep[for example, ][]{Jensen2001}.
Recently \cite{Sanchez-Franks2019} used a multi-year model simulation to argue that the ultimate source of the salty water is the western equatorial Indian Ocean.
Regardless of ultimate source, both observations and models agree that the Summer Monsoon Current (SMC) is the dominant pathway for salty water entering the Bay \citep{Jensen2001,Vinayachandran2013,Webber2018}\footnote{Recent observations and model simulations describe a second pathway — a persistent subsurface inflow of salty water during the NE monsoon that exists as a superposition of frequent salty intrusion events that average out to a region of broad northward flow of high salinity water west of 85°E \citep{Wijesekera2015,Jensen2016}.}.
The salty signature of the SMC is visible in maps of the depth of the 34.75 isohaline surface which shallows by \SI{25}{m} or so in the southwestern Bay during the summer monsoon (Figure \ref{fig:spatial}k--o; \citealp{Murty1992a,Vinayachandran2013}).
The shallow depth of the \(S = \SI{35}{psu}\) isohaline in the southwestern and south-central Bay relative to the northern Bay led \cite{Vinayachandran2013} to hypothesize that the southern Bay is a site of enhanced mixing and upward salt fluxes that may be an important contributor to the salt budget of the Bay.
In agreement with this hypothesis, model studies have implicated vertical mixing as the primary mechanism for diluting the immense amount of fresh water the Bay receives during the southwest monsoon \citep{Akhil2014,Benshila2014,Wilson2016a}.

@@

Figure \ref{fig:nrl} presents the seasonal cycle of winds, turbulence, shear and stratification at mooring NRL5 using daily averaged quantities.
We choose to highlight mooring NRL5  for two reasons.
First, it experiences the least blowdown and is least contaminated by the associated space-time aliasing (\SIrange{10}{20}{\meter}, Figure \ref{fig:nrl}f).
Second, the turbulence quantities in Figure \ref{fig:nrl} are inferred from measurements recorded by the deep χpod at 105m.
This instrument is the deepest deployed in the Bay to date, and recorded the longest period of
near-molecular diffusivity observed during the transition months of March and April.
The filtered shear components shown in Figure \ref{fig:nrl}d are obtained by first subsampling the filtered depth-time fields along the χpods trajectory and then normalizing by 30-day lowpass filtered \(N²\).
Time series recorded at the other moorings are presented in the Supplementary Material.

@@

We now describe the seasonal cycle of shear and turbulence by synthesizing Figures \ref{fig:shears}, \ref{fig:nrl} and \ref{fig:vert}.

@@

Our most dramatic observation is that the χpod at 104-m recorded near-laminar flow in the thermocline with near-molecular values of \(K_T\) during the \emph{entire} month of April.
Similar periods of low to negligible mixing are present at other χpods, but for shorter periods of time.
Median \(K_T ≤ \SI{1e-6}{\m\squared\per\s} \approxeq 5κ_T\) in most thermocline density bins (deeper distributions in Figure \ref{fig:vert}), so the observation of negligible mixing is consistent across all locations.
The transition months of March and April are a period of weak thermocline currents, weak thermocline shear, weak winds, high net surface heat flux and low near-inertial energy flux (Figures \ref{fig:spatial}, \ref{fig:shears} and \ref{fig:nrl}); this is consistent with the observations of low mixing.
Weak pulses of near-inertial shear are seen in Figures \ref{fig:shears} and \ref{fig:nrl}e; again this is consistent with weak wind forcing at the surface (Figure \ref{fig:spatial}k--o).
Note that stratification is relatively high; $N² \sim \SI{5e-4}{\per\second\squared}$ at all χpod depths.

@@

Elevated mixing at NRL5 occasionally lines up with short periods of elevated low frequency shear, indicating presence of the SMC, between May and October (Figure \ref{fig:nrl}e)
The SMC is visible in \(S\low\) at all three moorings during this season though for differing lengths of time (Figure \ref{fig:shears}).
Both seasonal mean surface velocities from the OSCAR product and mooring ADCP data show the SMC to be prominent especially at NRL3 and NRL4, the two westernmost moorings along 8°N (also see Figures \ref{fig:spatial}a--e, \ref{fig:hov} and \citealp{Wijesekera2016a}).
This inference is consistent with the ADCP measurements (Figure \ref{fig:shears}).

@@

There is a strong seasonal cycle in thermocline mixing (Figure \ref{fig:vert}) that appears to be linked to a seasonal cycle in thermocline shear (Figure \ref{fig:shears}).
The seasonal cycle in shear results from
(a) the seasonal presence of the Summer Monsoon Current which greatly increases low-frequency shear \(S\low\) between July and October, and
(b) episodic energetic downward propagating near-inertial waves observed outside March and April.
At times, \(S\low\) is of comparable magnitude to near-inertial shear \(S\niwp\) (Figure \ref{fig:shears}).
The seasonal cycle in low-frequency shear is expected from the well established seasonal spinup and spindown of the SMC \citep{Schott2001a}.
A seasonal cycle in near-inertial shear is perhaps expected from the seasonal cycle of winds but our ADCP record is not long enough to properly characterize the magnitude of the seasonal cycle in near-inertial energy.
Next we examine the consequences of turbulent mixing in the thermocline.

@@

Is observed seasonally-enhanced mixing in the south-central Bay's thermocline between May and November important for the Bay's salt budget?
The climatological depth of the \(S=\SI{34.75}{psu}\) surface at 8°N estimated using the Argo mapped climatology shallows by \SI{20}{m} or so between May and November relative to other months (Figures \ref{fig:spatial}k--o and \ref{fig:map}f--i).
The seasonal shallowing of this isohaline is particularly significant given that the observed diffusivity is surface intensified.
Mean diffusivity at this isohaline, thick orange horizontal line in Figure \ref{fig:vert}, is approximately \(\SI{1e-4}{\m\squared\per\s}\) during the SW monsoon and the post-monsoon period (SWNE) i.e. between May and November.
Diffusivity is an order of magnitude lower during the NE monsoon and near-molecular during the NESW transition.
Seasonally averaged surface velocities show the mean path of the SMC to be along the mooring line at 8°N (NRL3, NRL4, and NRL5; Figure \ref{fig:spatial}a--e), so we now attempt to quantify turbulent salt flux along 8°N in the south-central Bay using our admittedly sparse dataset.

@@

The sampling bias resulting from mooring blowdown suggests that we are likely underestimating the true magnitude of turbulent salt flux.
For example, all χpods at 8°N are forced down approximately \SI{50}{m} or so by the Summer Monsoon Current in July during which time they record little turbulent salt flux (Figure \ref{fig:8njs}a).
Inspection of the velocity fields shows that the χpods dive beneath the region of greatest shear in the water column and are likely missing the regions of greatest mixing during this period (Supplementary Information, Figures SXXXXXX \textbf{TODO}).
Given these uncertainties, we do not consider Figure \ref{fig:8njs}c a good estimate of the amplitude of the seasonal cycle of turbulent heat flux but instead interpret it as evidence that climatologically important turbulent fluxes occur in the south-central Bay at least between August and January.

@@

Year-long observations of turbulence from a moored mixing meter χpod at three locations along 8°N revealed a seasonal cycle in upper-ocean turbulence in the Bay of Bengal (Figures \ref{fig:map}, \ref{fig:vert} and Table \ref{tab:pods}).
The seasonal cycle of thermocline turbulence is influenced by the winds both through downward propagating near-inertial waves and by the strongly sheared Summer Monsoon Current (Figures \ref{fig:nrl}, \ref{fig:shears} and \ref{fig:nrl5-niw}).
Multiple χpods recorded extended periods of weak mixing (1--10 \(κ_T\)) between 50m and 100m depth during the months of March and April --- a period of weak winds, weak currents, weak shear and low near-inertial energy input (Figures \ref{fig:spatial}, \ref{fig:shears} and \ref{fig:nrl}; Tables B1,B2).
Despite these extended periods of low mixing, it has been hypothesized that turbulence in the Bay is necessary to close both heat and salt budgets \citep{Shenoi2002,Vinayachandran2013,Wilson2016a}.
Our observations suggest that turbulent salt fluxes of the right magnitude are indeed occurring in the south-central Bay (Section \ref{sec:results}\ref{sec:js}).

@@

Fully interpreting the observed seasonal cycle of mixing requires understanding the processes that drive and sustain the Bay's internal wave field.
We find that enhanced thermocline mixing generally coincides with observed bursts of near-inertial shear.
Understanding the many mechanisms and processes that drive the seasonal cycle of near-inertial shear at depth is thus of prime importance.
It is known that the strongly stratified transition layer at the base of the mixed layer can strongly influence the ability of winds to drive energy into the thermocline.
\cite{Dohan2011} studied observations during two different storms and found that in one case the wind-forced energy deepened the mixed layer with little to no mixing in the transition layer; while for a second storm of comparable magnitude, the mixed layer remain unchanged but transition layer was significantly broadened through mixing in the thermocline.
\cite{Brannigan2013} show that shear at the base of the transition layer depends on the alignment between ocean shear and wind stress.
Both studies imply that the near-surface fresh water layer could have a significant influence on the internal wave energy that ultimately leads to observed mixing.
\cite{Lucas2016} found this to be the case in the Bay --- they observed enhanced shear at the base of mixed layer but weak shear at the base of the barrier layer thereby isolating the thermocline from surface forcing.
Further complications arise from the interaction of near-inertial energy with lower-frequency mesoscale features in the Bay \citep{Johnston2016}.
Another related puzzle is the extended period of weak to negligible mixing during March and April.
This observation suggests that the Bay's internal wave field can be weaker than that expected from the Garrett-Munk spectrum, again highlighting the need for a study of the Bay's internal wave field.
The Bay thus offers intriguing opportunities for studying the ocean's internal wave field and its links to turbulence.
